---
title: "Duration Calculations"
author: "Laura F White"
date: "2023-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
source("utils.R")
library(readxl)
library(ggplot2)
library(reshape)
library(reshape2)
library(RColorBrewer)
library(ggpubr)
library(gridExtra)
library(grid)
library(scales)
library(dplyr)
library(data.table)
library(xlsx)
library(kableExtra)
library(tidyr)
```

```{r,include=F}
# read in data

# prevalence survey estimates, with years and CIs
dat.prev <- read_excel("Data/PrevalenceSurveyData.xlsx",sheet="Sheet1")

# data on treatments 
dat.cure <- read_excel("Data/AnnualTreatments.xlsx",sheet="Sheet1")

```

## Methods

### Estimation of the duration of disease
We leverage countries with two prevalence surveys to determine duration and incidence of TB each year between the two prevalence surveys. To do this, we also use data on cures and deaths from TB. We assume that the ratio of prevalence to incidence is not changing over the studied time period (which implies that disease duration is also not changing). As a first, most simple approach to this problem, we determine what the ratio of prevalence to incidence must be to get us to the value of the second prevalence survey. We calculate prevalence in year $t$, $P_t$ as a function of the prior year's prevalence, number treated, number who died and the incidence in the prior year, as follows: $P_t=P_{t-1}-C_{t-1}-D_{t-1}+I_{t-1}$, where $C_t$ and $D_t$ are the number of treated (or cured) cases of TB in year $t$ and $D_t$ are the number of deaths or self cures from TB in year $t$. We assume that the number who die or self cure have an average duration of 5 years and estimate the size of this population to be $D_t=(1/\delta_u)*(P_t-C_t)$, where $\delta_u$ is the duration of untreated disease. 

We observe prevalence at two time points and have annual treatment and mortality data.

We provide more detail on the estimates of prevalence for each country below. We couple prevalance per 100,000 individuals with the population level estimates when the prevalence surveys were conducted to get the total number of prevalent TB cases, which we use in our calculation. 

## Country specific Estimates

### Bangledesh

We do not have results for Bangladesh, yet. There is something odd that is producing really unusual results.

```{r, echo=FALSE}
country.tmp <- "Bangladesh"
bang.ann.rate <- getRateofChange(prev1=79,prev2=113,year1=2007,year2=2015)

bang.pop <- read.csv("Data/bang-pop.csv")

bang.dat <- getDuration2(country.tmp,pop.data=bang.pop)
bang.dat
```


### Vietnam

```{r, echo=FALSE}
country.tmp <- "Vietnam"
viet.ann.rate <- 1-0.045
viet.pop <- read.csv("Data/vietnam-pop.csv")

viet.dat <- getDuration2(country.tmp="Vietnam",pop.data=viet.pop)

col.names <- c("Country","Year","Prevalence","Treated","Incidence","Duration of treated disease",
               "Prevalence per 100k","Incidence per 100k","Assumed duration of untreated disease","Proportion of prevalence cases treated","Proportion of incident cases treated")
names(viet.dat) <- col.names

viet.dat %>% kbl() %>% kable_styling()
```


### Indonesia

```{r, echo=FALSE}
country.tmp <- "Indonesia"

indo.ann.rate <- getRateofChange(prev1=120,prev2=180,year1=2004,year2=2013.5)

indo.pop <- read.csv("Data/indo-pop.csv")

indo.dat <- getDuration2(country.tmp=country.tmp,pop.data=indo.pop)
names(indo.dat) <- col.names
indo.dat %>% kbl() %>% kable_styling()
```


### Philippines
 

```{r, echo=FALSE}
country.tmp <- "Philippines"
rate.change.10y <- 0.105
phil.prev1 <- dat.prev$`1st Prevalence`[dat.prev$Country==country.tmp] 
phil.ann.rate <- getRateofChange(prev1=phil.prev1,prev2=982,year1=2007,year2=2016)

phil.pop <- read.csv("Data/phil-pop.csv")

phil.dat <- getDuration2(country.tmp=country.tmp,pop.data=phil.pop)
names(phil.dat) <- col.names
phil.dat %>% kbl() %>% kable_styling()
```


### China

```{r, echo=FALSE}
country.tmp <- "China"
china.perc.change <- 0.35
china.prev1 <- dat.prev$`1st Prevalence`[dat.prev$Country==country.tmp]

china.ann.rate <- getRateofChange(prev1=china.prev1,prev2=108,year1=2000,year2=2010)

china.pop <- read.csv("Data/china-pop.csv")

china.dat <- getDuration2(country.tmp=country.tmp,pop.data=china.pop)
names(china.dat) <- col.names
china.dat %>% kbl() %>% kable_styling()
```



### Cambodia

```{r, echo=FALSE}
country.tmp <- "Cambodia"
cam.perc.change <- 0.452
cam.prev1 <- dat.prev$`1st Prevalence`[dat.prev$Country==country.tmp]

cam.ann.rate <- getRateofChange(prev1=cam.prev1,prev2=817,year1=2002,year2=2010)

cam.pop <- read.csv("Data/camb-pop.csv")

cam.dat <- getDuration2(country.tmp=country.tmp,pop.data=cam.pop)
names(cam.dat) <- col.names
cam.dat %>% kbl() %>% kable_styling()
```
 

### Myanmar

```{r, echo=FALSE}
country.tmp <- "Myanmar"
mya.perc.change <- 0.51
mya.prev1 <- dat.prev$`1st Prevalence`[dat.prev$Country==country.tmp]

mya.ann.rate <- getRateofChange(prev1=mya.prev1,prev2=415,year1=2009,year2=2017)

mya.pop <- read.csv("Data/myanmar-pop.csv")

mya.dat <- getDuration2(country.tmp=country.tmp,pop.data=mya.pop)
names(mya.dat) <- col.names
mya.dat %>% kbl() %>% kable_styling()
```
## Plotting the relationship between Duration and other parameters

```{r, echo=F}
# combnine all the results together
all.results <- dplyr::bind_rows(china.dat,cam.dat,mya.dat,phil.dat,indo.dat,viet.dat)

# plot duration by prevalence and incidence per 100k
plot.prev <- ggplot(all.results, aes(`Prevalence per 100k`,`Duration of treated disease`,color=`Year`)) +        
  geom_point(aes(shape=`Country`)) +
  theme_bw() 
plot.prev

plot.inc <- ggplot(all.results, aes(`Incidence per 100k`,`Duration of treated disease`,color=`Year`)) +        
  geom_point(aes(shape=`Country`)) +
  theme_bw() 
plot.inc

# plot with proportion of TB that is treated
plot.cure <- ggplot(all.results, aes(`Proportion of prevalence cases treated`,`Duration of treated disease`,color=`Year`)) +        
  geom_point(aes(shape=`Country`)) +
  theme_bw() 
plot.cure

plot.cure.inc <- ggplot(all.results, aes(`Proportion of incident cases treated`,`Duration of treated disease`,color=`Year`)) +        
  geom_point(aes(shape=`Country`)) +
  theme_bw() 
plot.cure.inc

```

## Sensitivity analysis for duration of untreated disease

We previously assumed that the duration of untreated disease is five years. We now vary this to be 4 years and 6 years to determine how much it alters the results.

### Vietnam

```{r, echo=FALSE}
country.tmp <- "Vietnam"
viet.ann.rate <- 1-0.045
viet.pop <- read.csv("Data/vietnam-pop.csv")

viet.dat.4 <- getDuration2(country.tmp="Vietnam",pop.data=viet.pop,unDxTBdur=4)
names(viet.dat.4) <- col.names

viet.dat.6 <- getDuration2(country.tmp="Vietnam",pop.data=viet.pop,unDxTBdur=6)
names(viet.dat.6) <- col.names

viet.all <- dplyr::bind_rows(viet.dat.4,viet.dat,viet.dat.6)

viet.all %>% kbl() %>% kable_styling()

# create plot of results
ggplot(viet.all,aes(as.factor(`Year`),`Prevalence per 100k`,color=as.factor(`Assumed duration of untreated disease`))) + 
  geom_point() +
  theme_bw() +
  xlab("Year") +
  labs(color="Assumed duration of untreated disease") +
  ggtitle("Vietnam")
```


### Indonesia

```{r, echo=FALSE}
country.tmp <- "Indonesia"

indo.ann.rate <- getRateofChange(prev1=120,prev2=180,year1=2004,year2=2013.5)

indo.pop <- read.csv("Data/indo-pop.csv")

indo.dat.4 <- getDuration2(country.tmp="Indonesia",pop.data=indo.pop,unDxTBdur=4)
names(indo.dat.4) <- col.names

indo.dat.6 <- getDuration2(country.tmp="Indonesia",pop.data=indo.pop,unDxTBdur=6)
names(indo.dat.6) <- col.names

indo.all <- dplyr::bind_rows(indo.dat.4,indo.dat,indo.dat.6)

indo.all %>% kbl() %>% kable_styling()

# create plot of results
ggplot(indo.all,aes(as.factor(`Year`),`Prevalence per 100k`,color=as.factor(`Assumed duration of untreated disease`))) + 
  geom_point() +
  theme_bw() +
  xlab("Year") +
  labs(color="Assumed duration of untreated disease") +
  ggtitle("Indonesia") +
  geom_point(aes(x=as.factor(2014),y=dat.prev$`2nd Survey Prevalence`[dat.prev$Country==country.tmp]),color="black",shape=3)
```


### Philippines
 

```{r, echo=FALSE}
country.tmp <- "Philippines"
rate.change.10y <- 0.105
phil.prev1 <- dat.prev$`1st Prevalence`[dat.prev$Country==country.tmp] 
phil.ann.rate <- getRateofChange(prev1=phil.prev1,prev2=982,year1=2007,year2=2016)

phil.pop <- read.csv("Data/phil-pop.csv")

phil.dat.4 <- getDuration2(country.tmp=country.tmp,pop.data=phil.pop,unDxTBdur=4)
names(phil.dat.4) <- col.names

phil.dat.6 <- getDuration2(country.tmp=country.tmp,pop.data=phil.pop,unDxTBdur=6)
names(phil.dat.6) <- col.names

phil.all <- dplyr::bind_rows(phil.dat.4,phil.dat,phil.dat.6)

phil.all %>% kbl() %>% kable_styling()

# create plot of results
ggplot(phil.all,aes(as.factor(`Year`),`Prevalence per 100k`,color=as.factor(`Assumed duration of untreated disease`))) + 
  geom_point() +
  theme_bw() +
  xlab("Year") +
  labs(color="Assumed duration of untreated disease") +
  ggtitle(country.tmp) +
  geom_point(aes(x=as.factor(2016),y=dat.prev$`2nd Survey Prevalence`[dat.prev$Country==country.tmp]),color="black",shape=3)
```


### China

```{r, echo=FALSE}
country.tmp <- "China"
china.perc.change <- 0.35
china.prev1 <- dat.prev$`1st Prevalence`[dat.prev$Country==country.tmp]

china.ann.rate <- getRateofChange(prev1=china.prev1,prev2=108,year1=2000,year2=2010)

china.pop <- read.csv("Data/china-pop.csv")

china.dat.4 <- getDuration2(country.tmp=country.tmp,pop.data=china.pop,unDxTBdur=4)
names(china.dat.4) <- col.names

china.dat.6 <- getDuration2(country.tmp=country.tmp,pop.data=china.pop,unDxTBdur=6)
names(china.dat.6) <- col.names

china.all <- dplyr::bind_rows(china.dat.4,china.dat,china.dat.6)

china.all %>% kbl() %>% kable_styling()

# create plot of results
ggplot(china.all,aes(as.factor(`Year`),`Prevalence per 100k`,color=as.factor(`Assumed duration of untreated disease`))) + 
  geom_point() +
  theme_bw() +
  xlab("Year") +
  labs(color="Assumed duration of untreated disease") +
  ggtitle(country.tmp) +
  geom_point(aes(x=as.factor(2010),y=dat.prev$`2nd Survey Prevalence`[dat.prev$Country==country.tmp]),color="black",shape=3)
```



### Cambodia

```{r, echo=FALSE}
country.tmp <- "Cambodia"
cam.perc.change <- 0.452
cam.prev1 <- dat.prev$`1st Prevalence`[dat.prev$Country==country.tmp]
y1 <- 2002
y2 <- 2011

cam.ann.rate <- getRateofChange(prev1=cam.prev1,prev2=817,year1=2002,year2=2010)

cam.pop <- read.csv("Data/camb-pop.csv")

cam.dat.4<- getDuration2(country.tmp=country.tmp,pop.data=cam.pop,unDxTBdur=4)
names(cam.dat.4) <- col.names

cam.dat.6 <- getDuration2(country.tmp=country.tmp,pop.data=cam.pop,unDxTBdur=6)
names(cam.dat.6) <- col.names

cam.all <- dplyr::bind_rows(cam.dat.4,cam.dat,cam.dat.6)

cam.all %>% kbl() %>% kable_styling()

# create plot of results
ggplot(cam.all,aes(as.factor(`Year`),`Prevalence per 100k`,color=as.factor(`Assumed duration of untreated disease`))) + 
  geom_point() +
  theme_bw() +
  xlab("Year") +
  labs(color="Assumed duration of untreated disease") +
  ggtitle(country.tmp) +
  geom_point(aes(x=as.factor(2011),y=dat.prev$`2nd Survey Prevalence`[dat.prev$Country==country.tmp]),color="black",shape=3)


```
 

### Myanmar

```{r, echo=FALSE}
country.tmp <- "Myanmar"
mya.perc.change <- 0.51
mya.prev1 <- dat.prev$`1st Prevalence`[dat.prev$Country==country.tmp]

mya.ann.rate <- getRateofChange(prev1=mya.prev1,prev2=415,year1=2009,year2=2017)

mya.pop <- read.csv("Data/myanmar-pop.csv")

mya.dat.4 <- getDuration2(country.tmp=country.tmp,pop.data=mya.pop,unDxTBdur=4)
names(mya.dat.4) <- col.names

mya.dat.6 <- getDuration2(country.tmp=country.tmp,pop.data=mya.pop,unDxTBdur=6)
names(mya.dat.6) <- col.names

mya.all <- dplyr::bind_rows(mya.dat.4,mya.dat,mya.dat.6)

mya.all %>% kbl() %>% kable_styling()

# create plot of results
ggplot(mya.all,aes(as.factor(`Year`),`Prevalence per 100k`,color=as.factor(`Assumed duration of untreated disease`))) + 
  geom_point() +
  theme_bw() +
  xlab("Year") +
  labs(color="Assumed duration of untreated disease") +
  ggtitle(country.tmp) +
  geom_point(aes(x=as.factor(2018),y=dat.prev$`2nd Survey Prevalence`[dat.prev$Country==country.tmp]),color="black",shape=3)


```

Finally we can assess what assumed duration of untreated disease allows us to get closest to the second prevalence survey estimate.

```{r, echo=F}
real.dat <- dat.prev %>% select(Country,`2nd Survey Prevalence`,`Second prevalence survey year`)

all.results <- dplyr::bind_rows(mya.all,china.all,cam.all,phil.all,viet.all,indo.all)

all.results.2 <- left_join(all.results,real.dat)

# has only the second prevalence survey results for each assumed duration
all.results.3 <- all.results.2 %>% filter(Year==`Second prevalence survey year`)

# now calculate the difference between the actual prevalence and estimated from the model
all.results.3$prev.diff <- abs(all.results.3$`2nd Survey Prevalence`-all.results.3$`Prevalence per 100k`)

# create nice table with the results we want
test <- spread(all.results.3[,c(1:2,9,12,14)], key = `Assumed duration of untreated disease`, value = `prev.diff`)
names(test) <- c(col.names[1:2],"2nd Survey Prevalence","Prev difference, d=4","Prev difference, d=5","Prev difference, d=6")
test2 <- spread(all.results.3[,c(1:2,6,9,12)], key = `Assumed duration of untreated disease`, value = `Duration of treated disease`)
names(test2) <- c(col.names[1:2],"2nd Survey Prevalence","Duration, d=4","Duration, d=5","Duration, d=6")
test3 <- spread(all.results.3[,c(1:2,7,9,12)], key = `Assumed duration of untreated disease`, value = `Prevalence per 100k`)
names(test3) <- c(col.names[1:2],"2nd Survey Prevalence","Prevalence, d=4","Prevalence, d=5","Prevalence, d=6")

summary.res.tmp <- left_join(test,test2)
summary.res <- left_join(summary.res.tmp,test3)

summary.res %>% kbl() %>% kable_styling()
```